---
title: "rGenomeTracks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rGenomeTracks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
rGenomeTracks package leverages the power of [pyGenomeTracks](https://github.com/deeptools/pyGenomeTracks) software
with the interactivity of R.
[pyGenomeTracks](https://github.com/deeptools/pyGenomeTracks) is a python software
that offers robust method for visualizing
epigenetic data files like narrowPeak, Hic matrix, TADs and arcs, however though,
here is no way currently to use it within R interactive session.
rGenomeTracks wrapped the whole functionality of pyGenomeTracks with additional utilites
to make to more pleasant for R users.

## Installing PyGenomeTracks
You should have pyGenomeTracks installed on your system. To avoid dependency clash, 
we highly recommed using miniconda. rGenomeTracks automatically install this dependency for you. You might be prompt 
to install miniconda on your system if not already installed later in this this example. 
If this step failed, you can manually install pyGenomeTracks either via reticulate or your commandline.


```{r setup}
# loading the rGenomeTracks
library(rGenomeTracks)
```

## Principle
rGenomeTracks deals creates tracks in a class `genome_track`. Currently, there are 14 tracks available:
1. track_bed()
2. track_bedgraph()
3. track_bedgraph_matrix()
4. track_gtf()
5. track_hlines()
6. track_spacer()
7. track_bigwig()
8. track_epilogos()
9. track_narrowPeak()
10. track_domains()
11. track_hic_matrix()
12. track_links()
13. track_scalebar()
14. track_x_axis()

Please refer to the help page for each one of them for details and examples.


```{r eval=FALSE}
# Download h5 example
 h5_dir <- tempfile(fileext = ".h5")
download.file("https://github.com/deeptools/pyGenomeTracks/raw/488a6d3f9e81c6ca1af9f90484445070ed9c7024/pygenometracks/tests/test_data/Li_et_al_2015.h5", destfile = h5_dir)
 
 # Load other examples
 tads_dir <- system.file("extdata", "tad_classification.bed", package = "rGenomeTracks")
 arcs_dir <- system.file("extdata", "links2.links", package = "rGenomeTracks")
 bw_dir <- system.file("extdata", "bigwig2_X_2.5e6_3.5e6.bw", package = "rGenomeTracks")
 
 # Create HiC track from HiC matrix
 h5 <- track_hic_matrix(
   file = h5_dir, depth = 250000, min_value = 5, max_value = 200,
   transform = "log1p", show_masked_bins = FALSE
 )

 # Create TADS track
 tads <- track_domains(
   file = tads_dir, border_color = "black",
   color = "none", height = 5,
   line_width = 5,
   show_data_range = FALSE,
   overlay_previous = "share-y"
 )

 # Create arcs track
 arcs <- track_links(
   file = arcs_dir, links_type = "triangles",
   line_style = "dashed",
   overlay_previous = "share-y",
   color = "darkred",
   line_width = 3,
   show_data_range = FALSE
 )

 # Create bigwig track
 bw <- track_bigwig(
   file = bw_dir, color = "red",
   max_value = 50,
   min_value = 0,
   height = 4,
   overlay_previous = "yes",
   show_data_range = FALSE
 )

```



`genome_track` objects can be added together using `+` function. 
 
```{r eval=FALSE}
 # Create one object from HiC, arcs and bigwid
 tracks <- h5 + arcs + bw
```


The track(s) to be plotted is to be passed to `plot_gtracks()`
for the generation of the plot. Additionally, `plot_gtracks()` requires the genomic region to be plotted. Optionally, you can set plot title, dpi, width, height, fontsize, track-to-label fraction, label alignment position, and directory to save the plot.
 
```{r eval=FALSE}
 # Plot the tracks
## Note to verify installing miniconda if not installed.
 plot_gtracks(tracks, chr = "X", start = 25 * 10^5, end = 31 * 10^5 )

 # Plot HiC, TADS and bigwig tracks
 plot_gtracks(h5 + tads + bw, chr = "X", start = 25 * 10^5, end = 31 * 10^5)

```

```{r echo=FALSE}
plot(imager::load.image(system.file("extdata", "images/example1.png", package = "rGenomeTracks")), axes = FALSE)
plot(imager::load.image(system.file("extdata", "images/example2.png", package = "rGenomeTracks")), axes = FALSE)
```
